---
title: 'STAT 22000: Homework 10'
author: "Zixi Li"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
mainfont: Times New Roman
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r setup, include=FALSE}
library(MASS)
library(GGally)
library(openintro)
library(mosaic)
library(knitr)
#library(tidyverse)
#library(ggformula)
#library(gridExtra)
library(broom)
options(width=70, digits=6, scipen=8)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
# Set R output size a bit smaller than default
opts_chunk$set(size='small') 
```
# Problem 1 Find the p-value
(a)
Upper one-sided p-value is 0.00771264
```{r}
pt(2.6,df=26-1,lower.tail = F)
```

(b)
Two-sided p-value is 0.0154253 
```{r}
2*pt(2.6,df=26-1,lower.tail = F)
```

(c)
Lower one-sided p-value is 0.0209623
```{r}
pt(-2.2,df=18-1)
```

(d)
Lower one-sided p-value is 0.979038
```{r}
pt(2.2,df=18-1)
```

# Problem 2 Compare different psychological therapies
# (a)
The sample mean \( \bar{x} \)=3

The standard deviation $s= 7.32042$
```{r}
weightchange=c(1.7, 11.7, -1.4, 0.7, 6.1, -0.8, -0.1, 1.1,
               2.4, -0.7, -4.0, 12.6, -3.5, 20.9,1.9, 14.9,
               -9.3, 3.9, 3.5, 2.1, 0.1, 17.1, 1.4, 15.4,
               -7.6, -0.3, -0.7,1.6, -3.7)
mean(weightchange)
sd(weightchange)
```

# (b)
$H_0$: $\mu=0$, $H_A$: $\mu>0$, where $\mu$ is the population mean weight gain of girls who received the cognitive behavioral therapy.

# (c)
$$df=n-1=29-1=28$$
$$t-statistic = \frac{\bar{x}-\mu_0}{s/\sqrt{n}}=\frac{3-0}{7.32/\sqrt{29}}=2.207$$

# (d)
The upper one-sided p-value is 0.0178404, which is larger than the significance level 0.01, so we fail to reject $H_0$, girls' mean weight gain is not significantly higher than 0.
```{r}
pt(2.207,df=28,lower.tail = F)
```

# (e)
Type 1 error: girls who receive the cognitive behavioral therapy have a 0 mean weight gain but we conclude that the mean weight gain is significantly higher than 0.

Type 2 error: girls who receive the cognitive behavioral therapy have mean wight gain higher than 0 but we fail to reject the null hypothesis.

# (f)
$H_0$: $\mu=0$, $H_A$: $\mu\neq0$, where $\mu$ is the population mean weight gain of girls who received the cognitive behavioral therapy.

The degree of freedom and t-statistic remain the same. df=28 and t-statistic=2.207

The two-sided p-value = $0.01784\times2=0.03568$

# (g)
The 95\% t-confidence interval is 2.04841, the corresponding mean weight gain is $3\pm 2.04841\times 7.32/\sqrt{29}=(0.2155, 5.7845)$, thus the true mean change in weight is positive, but possibly as small as 0.2155
```{r}
qt(0.05/2, df=28, lower.tail = F)
```

# (h)
We can see the t-statistic, the df, the p-values and the 95\% confidence level all agree with our calculation in (c)(d)(f)(g).
```{r}
cognitive = c(1.7, 11.7, -1.4, 0.7, 6.1, -0.8, -0.1, 1.1,
               2.4, -0.7, -4.0, 12.6, -3.5, 20.9,1.9, 14.9,
               -9.3, 3.9, 3.5, 2.1, 0.1, 17.1, 1.4, 15.4,
               -7.6, -0.3, -0.7,1.6, -3.7)
t.test(cognitive, alternative = "greater")
t.test(cognitive, alternative = "two.sided")
```

# Problem 3 True or faulse 
(a) is false. Increasing the significance level will increase the probability of making a Type 1 error.

(b) is true.

(c) is false. Failure to reject $H_0$ means we don't have strong evidence to reject it, we can't be sure that $\mu=5$. 


