---
title: "STAT 22000 Summer 2020 HW2 Solutions"
author: "Yibi Huang"
output: pdf_document
header-includes:
  - \usepackage[sol]{optional}
geometry: margin = .65in
fontsize: 11pt
---

```{r echo=F}
pdf.options(useDingbats = TRUE)
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, dpi=36)
```

\opt{hw}{
Beginners often have trouble loading data sets to R from a file.
Please refer to Section 3 of Lab \#1: \textcolor{blue}{http://www.stat.uchicago.edu/~yibi/s220/labs/lab01.html}
about how to change the working directory and load data set from a file.

You can refer to Lab \#2: 
\textcolor{blue}{http://www.stat.uchicago.edu/~yibi/s220/labs/lab02.html} to find
the details for R commands you need to make the histograms, boxplots, and scatterplots required below.


The data file `NLSY.txt` posted on Canvas along with HW2
contains data come from the 
NationalLongitudinal Study of Youth (NLSY), U.S. Bureau of Labor
Statistics https://www.bls.gov/nls/home.htm.
The subjects are 1306 American men and 1278 American women between
the ages of 14 and 22 in 1979. The variables include

* `Gender`
* `AQFT`: the percentile scores on the Armed Forces Qualifying Test,
which is designed for evaluating the suitability of military recruits
but which is also used by researchers as a general intelligence test
* `Edu2006`: years of education achieved by time of interview in 2006
* `Income2005`: annual income in thousands of dollars in 2005

The orignal data include more subjects but those with missing values of 
variables are omitted.
}

First, let's load the data set to R and load the mosaic library.

```{r loading-data-mosaic, message=FALSE}
NLSY = read.table("NLSY.txt",header=T)
library(mosaic)
```

# (a) 
\textcolor{red}{[2pts = 1pt for the histogram + 1pt for the comment on the modes of the histogram]}

As `Edu2006` is integer-valued, the natural bandwidth for the histogram is 1.

```{r 1a, cache=TRUE, fig.height=3}
histogram(~Edu2006, data=NLSY, width=1, xlab="Years of Education Completed by 2006")
```

The histogram has two peaks at 12 an 16 since most people either finished 
high school (12 years) or had a college degree (16 years).

# (b) 

\textcolor{red}{[2pts = 1pt for the plot and the summary + 1pt for the comment]}

From the histograms and the data summary below, 
the education level of males and females were pretty close.
They have identical five-number summaries 
and their means for `Edu2006` were pretty close 
(13.81 years for men and 13.91 years for women.)

```{r 1b, cache=TRUE, fig.height=4}
favstats(Edu2006 ~ Gender, data=NLSY)
histogram(~Edu2006 | Gender, data=NLSY, width=1, 
          xlab="Years of Education by 2006",layout=c(1,2))
```


# (c)

\textcolor{red}{[2pts = 1pt for the boxplot + 1pt for the missing info]}
The side-by-side boxplot is as follows.
We are not able to observe the modes at 12 and 16 from the boxplot which
can be observed in the histogram.

```{r 1c, cache=TRUE, fig.width=3, fig.height=2.5}
bwplot(Edu2006 ~ Gender, data=NLSY, ylab="Years of Education by 2006")
```


# (d) 
\textcolor{red}{[4pts in total = 1pt for the plots + 3pts for the comments (1pt each)]}

```{r 1d, cache=TRUE, fig.height=3.5, fig.show='hold', out.width=c('50%', '50%')}
histogram(~Income2005 | Gender, data=NLSY, width=10, 
          xlab="Annual Income in 2005 ($1000)",layout=c(1,2))
bwplot(Income2005 ~ Gender, data=NLSY, ylab="Annual Income in 2005 ($1000)")
favstats(Income2005 ~ Gender, data=NLSY)
```

* \textcolor{red}{[1pt]} The distributions of `Income2005` are __right-skewed__ for both gender
* \textcolor{red}{[1pt]} Males generally had a higher income.
The mean, Q1, median, and Q3 for males were all higher than those for females.
* \textcolor{red}{[1pt]} Males' distribution of income also had a higher variability than females'
as SD was 55.9 thousands of dollars for males and 28.8 thousands of dollars for 
females. One can also argue that males' histogram had a larger spread
or the height of the box (IQR) for males' boxplot is higher than that for females.


# (e) 

\textcolor{red}{[2pts = 1pt for the plots + 1pt for the left-skewness]}

After the log-transformation, the distribution of `Income2005` become 
__left-skewed__ (the left tail is longer).

```{r 1e, cache=TRUE, fig.width=4, fig.height=3.5, fig.show='hold', out.width=c('50%', '45%')}
histogram(~log(Income2005) | Gender, data=NLSY, width=0.5, 
          xlab="log of Annual Income in 2005 ($1000)",layout=c(1,2))
bwplot(log(Income2005) ~ Gender, data=NLSY, 
       ylab="log of Annual Income in 2005 ($1000)")
```


# (f)

\textcolor{red}{[2pts = 1pt for the plots + 1pt for the comparison]}

```{r 1f2, cache=TRUE, fig.width=8, fig.height=3.5}
NLSY$Edu2006.fac = factor(NLSY$Edu2006, labels = paste(6:20,"yr edu"))
bwplot(Income2005 ~ Gender | Edu2006.fac, data=NLSY, layout=c(8,2),
       ylab="Annual Income in 2005 ($1000)")
```

Taking logarithm might make it easier to tell which gender had a higher income.
We can see that men had a higher median income
than women in all the plots. 
Hence, we can conclude men earned more than women in general, 
even if they received the same number of years of education.

```{r 1f3, cache=TRUE, fig.width=8, fig.height=3.4}
bwplot(log(Income2005) ~ Gender | Edu2006.fac, data=NLSY,
      layout=c(8,2), ylab="log of Annual Income in 2005 ($1000)")
```


# (g)

\textcolor{red}{[4pts in total = 1pt for the plots + 3pts for the comments (1pt each)]}

1. \textcolor{red}{[1pt]} We can see that `Income2005`in general __increases__ with `AQFT`.
2. \textcolor{red}{[1pt]} The variability of `Income2005` __increases__ with `AQFT`. 
As `AQFT` gets higher, the spread of `Income2005` also gets higher.
3. \textcolor{red}{[1pt]} After taking logarithm, the variability of the __logarithm__ of `Income2005`
doesn't seem to change with `AQFT` 
as the points scatter in a band of constant width.

The scatter plots are shown below.

```{r 1g, cache=TRUE, fig.show='hold', out.width=c('50%', '50%')}
qplot(AFQT, Income2005, data=NLSY, shape=I(1),
      xlab="Percentile Score on the AFQT intelligence Test", 
      ylab="Annual Income in 2005 ($1000)")
qplot(AFQT, log(Income2005), data=NLSY, shape=I(1),
      xlab="Percentile Score on the AFQT intelligence Test",
      ylab="log of Annual Income in 2005 ($1000)")
```


# (h)

\textcolor{red}{[4pts in total = 1pt for the plots + 3pts for the comments (see below)]}


1. \textcolor{red}{[2pts]} From the plot below, the logarithm of 
`Income2005` increase with `AQFT` for each gender.
However, the variability of the logarithm of `Income2005`
doesn't seem th change with `AQFT` as the points just scatter in a band
of a constant width.

```{r 1h2, cache=TRUE, fig.width=3, fig.width=8}
qplot(AFQT, log(Income2005), data=NLSY,facets = ~Gender,
      xlab="Percentile Score on the AFQT intelligence Test",
      ylab="log of Annual Income in 2005 ($1000)")
```

2. \textcolor{red}{[1pt]} 
Comparing points with similar x-values in the plot below,
we see that blue points (males) tend to have higher y-values than 
points (females).
Men generally earned more than women, 
even if they have similar intellegence test score percentiles.

```{r 1h, fig.width=8, fig.height=3.5, cache=TRUE}
qplot(AFQT, log(Income2005), data=NLSY, color=Gender,
      xlab="Percentile Score on the AFQT intelligence Test",
      ylab="log of Annual Income in 2005 ($1000)")
```

<!-- 
__Remark__: (not required) The R code below added regression lines to the plot 
```{r 1g3} 
#qplot(AFQT, log(Income2005), data=NLSY, xlab="??", ylab="??",facets = ~Gender)+
    geom_smooth(method = "lm") 

#qplot(AFQT, log(Income2005), data=NLSY, xlab="??", ylab="??", color=Gender, shape=I(1))+
    geom_smooth(method = "lm") 
```
-->


# (i)

\textcolor{red}{[3pts]} In the split-panel scatter plot below, 
note that points in a single subplot
were of the same years of education.
We can compare the annual incomes of males and females
controlling for `Edu2006` and `AFQT`
by comparing the y-values of blue points (males) and red points (females)
with similar x-values in a single subplots as such
cases had identical `Edu2006` and similar `AFQT` percentiles.

We see that in nearly all of the subplots,
blue points (males) tend to have higher y-values than 
points (females).
We can then conclude that men 
generally earned more than women, 
even if they received the same years of education and had
the same intellegence test score percentiles.

```{r 1i, fig.width=8, fig.height=6, cache=TRUE}
qplot(AFQT, log(Income2005), data=NLSY,
      color=Gender, facets=~Edu2006.fac,
      xlab="Percentile Score on the AFQT intelligence Test",
      ylab="log of Annual Income in 2005 ($1000)")
```


(Not required.) Adding separate linear regression lines between `log(Income2005)` and `AFQT`
for men and women in each subplot makes it easier 
to compare the `log(Income2005)` of the two gender. 
(A linear regression line between x and y is the straight line that 
best depicts the linear relationship between x and y in a scatterplot.)
We can see that the blue lines are almost always above the red lines,
meaning that males generally had higher `log(Income2005)` than females,
after adjusting for `Edu2006` and `AFQT.`

```{r 1i2, fig.width=8, fig.height=6, cache=TRUE}
qplot(AFQT, log(Income2005), data=NLSY,
      color=Gender, facets=~Edu2006.fac, shape=I(1),
      xlab="Percentile Score on the AFQT intelligence Test",
      ylab="log of Annual Income in 2005 ($1000)")+
  geom_smooth(method="lm")
```


