---
title: "STAT 22000 2020 Summer HW3 Solutions"
author: "Yibi Huang"
output: pdf_document
geometry: margin = .75in
fontsize: 12pt
---

# 1. -- 9 total points

## (1a) 
\textcolor{red}{[1pt]} Let's first load the NC births data, 
exclude the case with missing `marital` status,
and load the `mosaic` library.
```{r message=FALSE}
nc = read.csv("https://www.openintro.org/stat/data/csv/ncbirths.csv")
nc = subset(nc, is.na(marital) == FALSE)
library(mosaic)
```

The two way table is

```{r}
tally(~ marital + lowbirthweight, data=nc)
```

The R command ``addmargins` adds the marginal totals to 
the two-way table.

```{r}
addmargins(tally(~ marital + lowbirthweight, data=nc))
```

## (1b) 
\textcolor{red}{[2pts in total, 1pt each]} 

About 
$\mathbf{386/999\approx0.3864\approx38.6\%}$ of the babies
were born to unmarried mothers.

About
$\mathbf{110/999\approx0.110=11\%}$ of the babies
had low birth weights.

Alternatively, the answers can be computed in R as follows.

```{r cache=TRUE}
prop.table(tally(~ marital, data=nc))
prop.table(tally(~ lowbirthweight, data=nc))
```


## (1c) 
\textcolor{red}{[1pt]} $\mathbf{61/386\approx0.158=15.8\%}.$ 
Alternatively, we can compute it using the R command:

```{r}
prop.table(tally(~ marital + lowbirthweight, data=nc),1)
```

## (1d) 
\textcolor{red}{[2pts]} The segmented bar chart is as follows.

```{r 1d, fig.width=3, fig.height=2.7, cache=TRUE}
barchart(tally(~ marital + lowbirthweight, data=nc), horizontal=FALSE)
```

(Not Required) It would be better to add legend and labels to the barplot(s).

```{r 1d1, fig.width=4, fig.height=2.5, cache=TRUE}
barchart(tally(~ marital + lowbirthweight, data=nc), 
         horizontal=FALSE,
         xlab="Mother",
         auto.key=list(space="right", title="Baby'sBirth Weight",cex.title=1))
```

##(1e) 
\textcolor{red}{[1pt]} 

```{r eval=FALSE, tidy=TRUE}
mosaicplot(tally(~ marital + lowbirthweight, data=nc), 
           ylab="Baby's Birth Weight", 
           xlab="Mother", color=TRUE,main="")
```

```{r 1e.mosaicplot, fig.width=3, fig.height=2.5, echo=FALSE, cache=TRUE}
par(mai=c(.56,.56,.1,.01),mgp=c(1.8,.5,0))
mosaicplot(tally(~ marital + lowbirthweight, data=nc), 
           ylab="Baby's Birth Weight", 
           xlab="Mother", color=TRUE, main="")
```

## (1f) 
\textcolor{red}{[1pt]} From the mosaic plot, 
we see babies born to unmarried women had
a higher percentage of having low birth weights
than those born to married women.

## (1g) 
\textcolor{red}{[1pt]} 
NOT independent because the locations of the segment in the two bars
do not match or because the unmarried group had 
a higher percentage of having low birth weight babies.

# 2. -- 10 total points

##(2a) 
\textcolor{red}{[3pts = 1pt for the table + 2pts for the proportions and the comparisons]}

The two way table with marginal total is

```{r}
addmargins(tally(~ marital + habit, data=nc))
```

About $\mathbf{60/613\approx0.098\approx9.8\%}$ 
  of the married mothers smoked during pregnancy,
and $\mathbf{66/386\approx0.171\approx17.1\%}$ 
of the unmarried mothers smoked during pregnancy.

Alternatively, we can compute using the R command:
```{r}
prop.table(tally(~ marital + habit, data=nc),1)
```
So the unmarried mothers were more likely to smoke during pregnancy.

## (2b)
```{r}
ftable(tally(~ habit + marital + lowbirthweight, data=nc))
prop.table(ftable(tally(~ habit + marital + lowbirthweight, data=nc)),1)
```

(b-i) \textcolor{red}{[2pts]} The percentages were 
$42/(42+511)\approx\fbox{7.6\%}$ for married nonsmoking moms, and $50/(50+270)\approx\fbox{15.6\%}$ unmarried nonsmoking moms.
Married nonsmoking moms were __less likely__ to have
low-birth-weight babies than unmarried nonsmoking moms.

(b-ii) \textcolor{red}{[2pts]} The percentages were 
$7/(7+53)\approx\fbox{11.7\%}$ for married smoking moms, and $11/(11+55)\approx\fbox{16.7\%}$ for unmarried smoking moms.
Married smoking women were __less likely__ to have
low-birth-weight babies than unmarried smoking women.

(b-iii) \textcolor{red}{[1pt]} After adjusting for smoking status,
unmariried women still had higher rates of 
having low-birth-weight babies than married women.
Out-of-marriage births are still associated with low-birth-weights,
even after accounting for the effect of smoking.

## (2c) 
\textcolor{red}{[2pts]} No, this is an observational study.
Even if we tried to rule out one confounding factor of smoking,
there can be other confounders like mothers' age, 
income, or education level that may be 
associated with mothers' marital status
and how likely babies may have low birth weights.
We cannot make a causal conclusion here.

# 3. -- 6 total points

## (3a) 
\textcolor{red}{[2pts]} Yes. Since subjects were __randomized__ to receive either
bupropion or a placebo.

## (3b) 
\textcolor{red}{[2pts]} Yes, the study blocked on whether a subject lived with a smoker.

## (3c) 
\textcolor{red}{[2pts]} No, since the subjects were __not randomized__ to
live with a smoker or not to live with a smoker. 
The comparison between those who lived with a smoker and those
who didn't is observational and is hence subject to confounding.