---
title: 'STAT 22000: Final Part 3 '
author: "Zixi Li"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r setup, include=FALSE, message =FALSE}
#library(MASS)
#library(GGally)
#library(openintro)
library(mosaic)
library(knitr)
#library(tidyverse)
#library(ggformula)
#library(gridExtra)
library(broom)
options(width=70, digits=6, scipen=8)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
# Set R output size a bit smaller than default
opts_chunk$set(size='small') 
# Set the default to NOT display R code
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=3.5)
nc = read.csv("https://www.openintro.org/stat/data/csv/ncbirths.csv")
```

## Q1 
With the following R codes, we know that

* the number of cases that the mother is married: 613
* the number of cases that the mother is unmarried: 386
* the averae age of the married mothers in the data: 29.14
* the averae age of the unmarried mothers in the data: 23.56
* the standard deviation (SD) of the age of the married mothers in the data: 5.52
* the standard deviation (SD) of the age of the unmarried mothers in the data: 5.66

```{r Q1}
favstats(mage ~ marital, data=nc)
```

## Q2
With $df=min(613-1,386-1)=385$, the critical value for 95\% CI is $t^{*} \approx 1.966$.

```{r Q2}
qt(0.05/2, df=385, lower.tail = F)
```

The 95\% CI for the mean difference $\mu_m-\mu_n$ is

$$\bar{x}_{m}-\bar{x}_{n} \pm t^{*} \sqrt{\frac{s_{m}^{2}}{n_{m}}+\frac{s_{n}^{2}}{n_{n}}} \approx 29.14-23.56 \pm 1.966 \sqrt{\frac{(5.52)^{2}}{613}+\frac{(5.66)^{2}}{386}}\approx (4.864,6.296)$$
Here the subscript "m" represents married, "n" represents not married.

## Q3
The age of married mothers is independent with the age of unmarried mothers.

The sample size $n_m=613$ and $n_n=386$ are both sufficiently large so that the t-approximation is good even when the data are clearly skewed. 




