---
title: 'STAT 22000: Homework 11'
author: "Zixi Li"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
mainfont: Times New Roman
header-includes: \usepackage{pdfpages}
geometry: margin=0.75in
fig_crop: no
---

```{r setup, include=FALSE}
library(MASS)
library(GGally)
library(openintro)
library(mosaic)
library(knitr)
#library(tidyverse)
#library(ggformula)
#library(gridExtra)
library(broom)
options(width=70, digits=6, scipen=8)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
# Set R output size a bit smaller than default
opts_chunk$set(size='small') 
```
# Problem 1 Confidence interval and significance level
(a) $H_o$ will be rejected at significance level $\alpha=0.05$, because $\alpha=0.05$ corresponds to a 95\% confidence level, and $\mu=3$ is not in the 95\% confidence level.

(b) $H_o$ will not be rejected at significance level $\alpha=0.01$, because $\alpha=0.01$ corresponds to a 99\% confidence level, and $\mu=3$ is in the 99\% confidence level.

# Problem 2 P-value and confidence interval
(a) $\mu=10$ is not included in the 95\% confidence interval, because 95\% confidence interval corresponds to 0.05 significance level, and p-value 0.03 will make $H_0$ be rejected at 0.05 significance level.

(b) $\mu=10$ is included in the 99\% confidence interval, because 99\% confidence interval corresponds to 0.01 significance level, and p-value 0.03 will make $H_0$ fail to be rejected at 0.01 significance level.

# Problem 3 Allergies and antigens levels
# (a)
```{r fig.width=5, fig.height=3, fig.align='center'}
Endotoxin = c(708.23, 911.60, 976.81, 1316.63, 262.74, 9772.08,
              370.76, 229.16, 2570.51, 891.19, 3163.20, 1777.65,
              1288.57, 436.23, 2631.63, 1173.52, 911.67, 7942.42,
              740.32, 356.92, 1175.48, 1480.55, 2754.61, 575.62,
              573.89, 468.26, 1000.71, 364.22, 1025.26, 1022.04,
              645.41, 363.57, 977.47, 1022.75, 1860.63, 371.13,
              174.73, 399.68, 1479.77, 2882.96, 601.99, 1697.32,
              2291.00, 646.49, 1176.27, 1995.43, 955.54, 1480.05,
              456.71, 1174.70, 5494.22) 
bwplot(Endotoxin, horizontal=T) 
```

Based on the above boxplot, we can see there are some outliers, and the shape is right-skewed, so it's not appropriate to construct a t-conﬁdence interval for the mean endotoxin level at the homes of children without allergy.

# (b)
```{r fig.width=5, fig.height=3, fig.align='center'}
bwplot(log(Endotoxin), horizontal=T)
```

Based on the above boxplot, we can see there are only two outliers, and the shape is not skewed, so it's more appropriate to construct a t conﬁdence interval for the mean of the log endotoxin level at the homes of children without allergy. 


# (c)
We can find the sample mean \( \bar{x} \)=6.91658 and the sample SD $s=0.858441$ and the critical value $t^*= 2.00856$ using the following R codes:
```{r}
mean(log(Endotoxin))
sd(log(Endotoxin))
qt(0.05/2,df=51-1,lower.tail = F)
```

Then the 95\% t-confidence interval can be calculated as:
$$\bar{x}\pm t^* \frac{s}{\sqrt{n}}=6.91658 \pm 2.00856 \times \frac{0.858441}{\sqrt{51}}=(6.67514, 7.15802)$$

# Problem 4 Measurements of body dimensions
# (a)
Here I make a histogram for the height of the 260 women and calculate the population mean and SD using the following R codes:
```{r fig.width=5, fig.height=3, fig.align='center' }
download.file("http://www.openintro.org/stat/data/bdims.RData",
              destfile = "bdims.RData")
load("bdims.RData")
fdims = subset(bdims, sex == 0)
population = fdims$hgt
mu = mean(population); mu
sigma = sd(population); sigma
histogram(population, fit="normal", nint=15, xlab="Women's height (cm)")
```

We can see that the population mean $\mu=164.872$, the SD $\sigma= 6.5446$. From the shape of histogram we know there's no obvious outlier nor clear skewness.

# (b) (1)
Here I plot the 90\% z-CI with known population SD $\sigma$: \( \bar{x} \) $\pm 1.645\sigma/\sqrt{n}$ with following codes:
```{r include=FALSE}
panel.ci = function (x, y, mu, data2, ...) {
  panel.xyplot(x, y, ...)
  good = subset(data2, upper > mu & lower < mu)
  bad = subset(data2, upper < mu | lower > mu)
  with(good, panel.arrows(lower, .index, upper, .index, angle=90, length = 0.05, ends="both"))
  with(bad, panel.arrows(lower, .index, upper, .index, angle=90, length = 0.05, ends="both", col="red", lwd=3))
  panel.abline(v=mu, lty=2)
}

plot_ci = function (df, mu, ...) {  
  space = 0.1 * (max(df$lower) - min(df$upper))
  xyplot(.index ~ mean, data=df, data2 = df, panel=panel.ci, mu = mu, xlim = c(min(df$lower) - space, max(df$upper) + space), ylab = NULL)
}
```

```{r fig.width=6, fig.height=4, fig.align='center'}
samp = do(100)*favstats(sample(population, size=5))
sigma=sd(population)
samp=transform(samp, lower=mean-1.645*sigma/sqrt(5))
samp=transform(samp, upper=mean+1.645*sigma/sqrt(5))
plot_ci(samp, mu=mean(population))
```

For the 100 CIs constructed, we notice 11 of them have missed the true population mean, the proportion of intervals that include the true population mean is 89\%, which is close to nominal level 90\%.

# (b) (2)
Here I plot the 90\% z-CI with unknown population SD: \( \bar{x} \) $\pm 1.645 s/\sqrt{n}$ with following codes:
```{r fig.width=6, fig.height=4, fig.align='center'}
samp=transform(samp, lower=mean-1.645*sd/sqrt(5))
samp=transform(samp, upper=mean+1.645*sd/sqrt(5))
plot_ci(samp, mu=mean(population))
```

For the 100 CIs constructed, we notice 18 of them have missed the true population mean, the proportion of intervals that include the true population mean is 82\%, which is lower than nominal level 90\%.

# (b) (3)
Here I plot the 90\% t-intervals:\( \bar{x} \) $\pm t^* s/\sqrt{n}$ with following codes, where $t^*=2.13185$
```{r fig.width=6, fig.height=4, fig.align='center'}
qt(0.1/2, df=4, lower.tail = F)
samp=transform(samp, lower=mean-2.13185*sd/sqrt(5))
samp=transform(samp, upper=mean+2.13185*sd/sqrt(5))
plot_ci(samp, mu=mean(population))
```

For the 100 CIs constructed, we notice 9 of them have missed the true population mean, the proportion of intervals that include the true population mean is 91\%, which is close to nominal level 90\%.